%% ----------------------------------------------------------------
%% Thesis.tex -- MAIN FILE (the one that you compile with LaTeX)
%% ---------------------------------------------------------------- 

% Set up the document
\documentclass[a4paper, oneside, 11pt, hidelinks]{Thesis}  % Use the "Thesis" style, based on the ECS Thesis style by Steve Gunn
\graphicspath{Figures/}  % Location of the graphics files (set up for graphics to be in PDF format)
\usepackage[T1]{fontenc}
\usepackage{helvet}
\renewcommand{\familydefault}{\sfdefault}
% Include the extra LaTeX packages required
\usepackage[square, numbers, comma, sort]{natbib}  % Use the "Natbib" style for the references in the text
\usepackage{verbatim}  % Needed for the "comment" environment to make LaTeX comments
\usepackage{vector}  % Allows "\bvec{}" and "\buvec{}" for "blackboard" style bold vectors in maths
\hypersetup{urlcolor=black, colorlinks=false}  
\usepackage{caption} 
\usepackage{subcaption} % This relies on commenting out line 111 in Thesis.cls to remove "subfigure"
\usepackage{datetime}
\usepackage{lscape}
\usepackage{array} % for the \newcolumntype facility used for our custom column type
\usepackage{multirow} % to allow spanning rows and columns in tables
\usepackage[a-1b]{pdfx} % For PDF/A generation
\usepackage{pdfpages} % To allow inclusion of PDF files
\usepackage{xcolor} % Just to allow first footnote to be subtle grey

% Suppress warnings about Underfull or Overfull boxes
\hbadness=99999

% This allows code formatting blocks through the 'listings' package
\usepackage{listings}
\lstset{frame=tb,
  language=Java,
  aboveskip=6mm,
  belowskip=6mm,
  frame=none,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=none,
  % numberstyle=\tiny\color{gray},
  % keywordstyle=\color{blue},
  % commentstyle=\color{dkgreen},
  % stringstyle=\color{mauve},
  breaklines=true,
  breakatwhitespace=true,
  tabsize=3
}

\newcolumntype{P}[1]{>{\centering\arraybackslash\hspace{0pt}}p{#1}}

%% ----------------------------------------------------------------
\begin{document}
\frontmatter      % Begin Roman style (i, ii, iii, iv...) page numbering

% Set up the Title Page
\title  {Addressing Energy Efficiency in System Design: A Journey from Architecture to Operation}
\authors  {\texorpdfstring
            {Eoin Woods}
            {Eoin Woods}
            }
\addresses  {\groupname\\\deptname\\\univname}  % Do not change this here, instead these must be set in the "Thesis.cls" file, please look through it instead
\date       {December 2018}
\subject    {}
\keywords   {}

\maketitle
%% ----------------------------------------------------------------

\pagestyle{plain}  % Set the page headers to the "fancy" style

% The Abstract Page
\addtotoc{Abstract}  % Add the "Abstract" page entry to the Contents
\abstract{
\setstretch{1.1}  % Set to small inter-line spacing to get abstract on one page

Digital-transformation initiatives have led to major efficiencies and cost savings but at the cost of consuming nearly 10 percent of the world's electricity.  Energy consumption research has increased datacentre, network, and hardware efficiency, but a neglected aspect of energy research has been the energy consumption of the software applications that underpin digital transformation.  To date, software architects have lacked the knowledge, guidance, and tools to allow them to understand the energy properties of their systems.

The research reported in this thesis begins to address this situation by developing practical knowledge, techniques, and tools to allow software architects to play their part in controlling the energy consumption of our modern digital world.

The work commences with an investigation into formal architectural description languages, through a literature review and a case study, resulting in two research contributions, namely a comprehensive systematic survey of architecture description languages from 1991 to 2015, and a case study of practical ADL use at scale in industry.

The second part of the research investigates how to assist architects in prioritising energy efficiency through a study of how experienced architects focus their attention for maximum effectiveness, which leads to the development of a model to guide architecture practitioners, which is validated and refined through a large survey of practising software architects.  The research contribution is a refined and validated model for architectural effort prioritisation.

The third aspect of the research examines the energy-related guidance available to architects and having found little generally applicable advice, analyses a significant industrial case study to understand how leading-edge practitioners addressed energy efficiency, contributing a set of three energy-related architectural principles, which can be used to guide architects in improving application energy efficiency.

Finally, we consider the practical problem of understanding the runtime energy properties of a system, and designed a novel approach to estimate the energy consumption of execution scenarios via application execution tracing and a cost-based energy model.  We created a proof of concept implementation of the approach and validated its consistency and correctness through practical testing.  The contribution of this work was twofold, namely the design of a practical system for allocating energy to application execution scenarios, and a tested, open-source, proof-of-concept implementation of the system.

Hence, the result of this work is six distinct contributions to knowledge in the area of ADLs (the survey and practical case study), architectural practice (the prioritisation model and the architectural principles for energy efficiency) and application energy efficiency (the design of the energy allocation system and the proof-of-concept implementation), which collectively can help architects to treat energy efficiency as a first class architectural concern in their work

\begin{tiny}
{\setlength{\parindent}{0cm}
	\setlength{\parskip}{0em}
	
	\textcolor{lightgray}{20181218.3}
}
\end{tiny}

\clearpage  % Abstract ended, start a new page
%% ----------------------------------------------------------------

% Declaration Page required for the Thesis, your institution may give you a different text to place here
%\Declaration{
%
%\addtocontents{toc}{\vspace{1em}}  % Add a gap in the Contents, for aesthetics
%
%I, AUTHOR NAME, declare that this thesis titled, `THESIS TITLE' and the work presented in it are my own. I confirm that:
%
%\begin{itemize} 
%\item[\tiny{$\blacksquare$}] This work was done wholly or mainly while in candidature for a research degree at this University.
% 
%\item[\tiny{$\blacksquare$}] Where any part of this thesis has previously been submitted for a degree or any other qualification at this University or any other institution, this has been clearly stated.
% 
%\item[\tiny{$\blacksquare$}] Where I have consulted the published work of others, this is always clearly attributed.
% 
%\item[\tiny{$\blacksquare$}] Where I have quoted from the work of others, the source is always given. With the exception of such quotations, this thesis is entirely my own work.
% 
%\item[\tiny{$\blacksquare$}] I have acknowledged all main sources of help.
% 
%\item[\tiny{$\blacksquare$}] Where the thesis is based on work done by myself jointly with others, I have made clear exactly what was done by others and what I have contributed myself.
%\\
%\end{itemize}
% 
% 
%Signed:\\
%\rule[1em]{25em}{0.5pt}  % This prints a line for the signature
% 
%Date:\\
%\rule[1em]{25em}{0.5pt}  % This prints a line to write the date
%}
%\clearpage  % Declaration ended, now start a new page

%% ----------------------------------------------------------------
\addtotoc{Contents}
\lhead{\emph{Contents}}  % Set the left side page header to "Contents"
\tableofcontents  % Write out the Table of Contents

%% ----------------------------------------------------------------
\lhead{\emph{List of Figures}}  % Set the left side page header to "List if Figures"
\listoffigures  % Write out the List of Figures

%% ----------------------------------------------------------------
\lhead{\emph{List of Tables}}  % Set the left side page header to "List of Tables"
\listoftables  % Write out the List of Tables

%% ----------------------------------------------------------------
% The Acknowledgements page, for thanking everyone
\acknowledgements{
%\addtocontents{toc}{\vspace{1em}}  % Add a gap in the Contents, for aesthetics

Every PhD student thanks their supervisor in their acknowledgements but in this case it is particularly important.  Without the guidance, support, inspiration and friendship of Dr Rabih Bashroush, this research would not have been started, let alone completed.  Thank you, Rabih.  You've been a strong, calm and knowledgeable guide through this complex, and at times intimidating, process.  I  wouldn't have completed it without you.

I'd also like to thank the many members of UEL's Computer Science and Informatics group who have encouraged, advised or guided me during the process including my second supervisor Dr Syed Islam, Professor Allan Brimicombe, Dr Andres Baravalle, Dr Aloysius Edoh, Dr Paolo Falcarin and Mr Mike Kretsis.  Charlotte Forbes has also provided invaluable guidance and administrative support for graduate students in the department.

The other group of people who have knowingly or unknowingly inspired, supported or encouraged me during this long process are the software architecture research community, particularly the communities who attend the ECSA and WICSA conferences.  I fear that I will undoubtedly forget someone, but I am most grateful for the friendship, guidance and support at different times of this group, including Professor Philippe Kruchten, whose work has inspired me throughout my career, Dr Eltjo Poort of CGI whose own long journey through a part-time PhD encouraged me to keep progressing mine, Dr John Klein, Dr Grace Lewis, Dr Rod Nord and Dr Ipek Ozkaya of the SEI who have encouraged me repeatedly along the way, Professors Hans van Vliet and Patricia Lago from Vrije Universiteit, Amsterdam and Dr George Fairbanks of Google and Michael Keeling from IBM who always inspire me to think more deeply about things that I think I understand.

I'd also like to thank my colleagues from Endava, who create a collegiate, friendly and supportive environment in which to work, particularly my managers John Cotterell our CEO, Rob Machin our COO, and all of my colleagues in our delivery units, and our strategy and operations groups.

Finally, but most importantly, I'd like to thank my family who have encouraged me throughout, despite never being quite clear what I'm doing, or why on earth I am doing a PhD in my 40s.  Thank you, Lynda and Katherine, for putting up with the long periods while I've been stuck in my study during evenings and weekends.  Thank you, Mum and Dad, for your continual support and belief in me.  Thank you, Nick, Marcia, Shiann and Kallum, for all you bring our family and reminding me to step away from the research sometimes when more important things need some attention!
\newline
\newline
\newline
Eoin Woods \newline
September 2018

}
\clearpage  % End of the Acknowledgements

%% ----------------------------------------------------------------
% End of the pre-able, contents and lists of things
% Begin the Dedication page
\pagestyle{empty}  % Page style needs to be empty for this page
\dedicatory{I dedicate this work to my family.  My wife Lynda and daughter Katherine, who make it all worthwhile.  My parents, Desmond and Anne, who have supported me in so many ways and to whom I owe so much.  And Nick, Marcia, Shiann and Kalum, who bring so much love and light to our family.}

\addtocontents{toc}{\vspace{2em}}  % Add a gap in the Contents, for aesthetics


%% ----------------------------------------------------------------
\setstretch{1.5}  % Return the line spacing back to 1.5
\mainmatter	  % Begin normal, numeric (1,2,3...) page numbering
\pagestyle{plain}  % Return the page headers back to the "fancy" style

% Include the chapters of the thesis, as separate files
% Just uncomment the lines as you write the chapters

\input{Chapters/introduction-chapter} %Ch1

\input{Chapters/literature-review-chapter} %Ch2

\input{Chapters/using-adls-chapter} % Ch3 - ADLs & Case Study

\input{Chapters/prioritisation-chapter} % Ch4  - Focusing Architectural Attention

\input{Chapters/energy-efficient-design-principles-chapter} %Ch5 - Design Principles for Energy Efficient Applications

\input{Chapters/monitoring-application-energy-chapter} %Ch6 - Monitoring Application Energy Usage During Operation

\input{Chapters/implementation-of-apollo-chapter} %Ch7 - Implementation of Application Energy Monitoring

\input{Chapters/apollo-validation-chapter} %Ch8 - Validation of Application Energy Monitoring

\input{Chapters/conclusions-chapter} % Ch9

%% ----------------------------------------------------------------
% Now begin the Appendices, including them as separate files

%\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics

\appendix % Cue to tell LaTeX that the following 'chapters' are Appendices

\input{Appendices/adl-list-appendix}
\input{Appendices/ethics-approval-appendix}

%\input{Appendices/AppendixB} % Appendix Title

%\input{Appendices/AppendixC} % Appendix Title

%\addtocontents{toc}{\vspace{2em}}  % Add a gap in the Contents, for aesthetics
\backmatter

%% ----------------------------------------------------------------
\raggedright
\label{Bibliography}
\lhead{\emph{Bibliography}}  % Change the left side page header to "Bibliography"
\bibliographystyle{plain}  % Use the "plain" BibTeX style for formatting the Bibliography
\bibliography{Bibliography,WebBibliography}  % The references are stored in the files "Bibliography.bib" and "WebBibliography.bib"

\end{document}  % The End
%% ----------------------------------------------------------------
